Perturbing SU($n$)-symmetric interactions

### 2019-11-15 collective_sun/sun_perturbation.tex
added second-order results for ZZ interactions, and some second-order results for a general two-body perturbation

We now return to the task of computing the second-order effective
Hamiltonian $H_1^{(2)}$.  Any coefficient vector $\v v$ can be
expanded into its projections $\v v_\Delta$ onto the eigenspace of
vectors satisfying \eqref{eq:cond_1} with eigenvalue $\Delta$,
i.e.~$\v v = \sum_\Delta \v v_\Delta$, which also allows us to expand
\begin{align}
  \V_1\p{\v v} = \sum_\Delta \V_1\p{\v v_\Delta},
\end{align}
where each operator $\V_1\p{\v v_\Delta}$ generates a state with
definite interaction energy $\Delta$ above that of the fully symmetric
manifold $\M_0$.  We can therefore simplify
\begin{align}
  H_1^{(2)}
  = - \P_0 \V_1\p{\v v} \E \V_1\p{\v v} \P_0
  = - \sum_{\Delta>0} \f1{\Delta}
  \P_0 \V_1\p{\v v} \P_\Delta \V_1\p{\v v} \P_0
  = - \sum_{\Delta>0} \f1{\Delta} \P_0 \V_1\p{\v v_\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \V_1\p{\v v_\Delta}^2 \P_0$ is worked out in
Appendix \ref{sec:PXXP}.  In total, we find
\begin{align}
  H_1^{(2)} \EQFS \tilde v \sp{\p{\col{X}}^2 - N \p{\col{X^2}}},
  \label{eq:H_1_2}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, and
\begin{align}
  \tilde v \equiv \f1{N\p{N-1}}
  \sum_{\Delta>0} \f{\v v_\Delta\c\v v_\Delta}{\Delta},
  &&
  \col{X^2} \equiv \sum_p X_p^2.
\end{align}
As the vectors $\v v_\Delta$ with $\Delta>0$ are mean-zero, we can
equivalently write
\begin{align}
  \tilde v = \f1{N-1}
  \sum_{\Delta>0} \f{\var\p{\v v_\Delta}}{\Delta},
\end{align}
where $\var\p{\v x}$ is the variance of the entries in $\v x$.

@@

If we {\it construct} a perturbation $\V_2\p{\m w_\Delta}$ with
coefficients $\m w_\Delta$ that satisfy \eqref{eq:cond_2} or
\eqref{eq:cond_2_eig} with eigenvalue $\Delta/2$, then
\begin{align}
  H_0 \V_2\p{\m w_\Delta} \ket\psi
  = \p{E_0 + \Delta} \V_2\p{\m w_\Delta} \ket\psi,
\end{align}
which implies that $\V_2\p{\m w_\Delta}\ket\psi$ is a state of
definite energy that depends only on the coefficients $\m w_\Delta$.
As with the case of single-body perturbations, we can decompose any
matrix $\m w$ into its projections $\m w_\Delta$ onto the eigenspace
of matrices satisfying \eqref{eq:cond_2} with eigenvalue $\Delta/2$,
i.e.~$\m w=\sum_\Delta\m w_\Delta$, which allows us to expand
\begin{align}
  \V_2\p{\m w} = \sum_\Delta \V_2\p{\m w_\Delta},
\end{align}
and in turn simplify
\begin{align}
  H_2^{(2)} = - \P_0 \V_2\p{\m w} \E \V_2\p{\m w} \P_0
  = -\sum_{\Delta>0} \f1\Delta \P_0 \V_2\p{\m w_\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \V_2\p{\m w_\Delta}^2 \P_0$ can be written in
terms collective operators using techniques similar to those in
Appendix \ref{sec:sym_prod}.  This calculation is partially completed
in Appendix \ref{sec:PXYXYP}.

@@

Calculating the second-order effective Hamiltonian $H_1^{(2)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_1$ in
\eqref{eq:perturbations} requires us to simplify
\begin{align}
  \P_0 \V_1\p{\v v}^2 \P_0
  \EQFS \sum_{p,q} v_p v_q X_p X_q
  = \sum_p v_p^2 X_p^2
  + \sum_{p\ne q} v_p v_q X_p X_q,
  \label{eq:PXXP_start}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$; and the coefficients $\v v$ satisfy
\eqref{eq:cond_1} with eigenvalue $\Delta\ne0$, which implies that
$\v v$ is mean-zero.  The first sum on the right of
\eqref{eq:PXXP_start} is then
\begin{align}
  \sum_p v_p^2 X_p^2
  \EQFS \sum_p v_p^2 X_0^2
  \EQFS \f1N \v v\c\v v\, \col{X^2},
  &&
  \col{X^2} \equiv \sum_p X_p^2.
  \label{eq:PXXP_eq}
\end{align}
The second sum on the right of \eqref{eq:PXXP_start} is
\begin{align}
  \sum_{p\ne q} v_p v_q X_p X_q
  \EQFS \sum_{p\ne q} v_p v_q X_0 X_1,
  \label{eq:PXXP_neq_start}
\end{align}
where the sum over coefficients in \eqref{eq:PXXP_neq_start} can be
simplified to
\begin{align}
  \sum_{p\ne q} v_p v_q
  = \sum_{p,q} v_p v_q - \sum_{p=q} v_p v_q
  = - \v v\c\v v
\end{align}
In order to simplify the product of operators in
\eqref{eq:PXXP_neq_start}, we expand
\begin{align}
  \p{\col{X}}^2
  = \sum_p X_p^2 + \sum_{p\ne q} X_p X_q
  \EQFS \col{X^2} + N \p{N-1} X_0 X_1,
  \label{eq:PXXP_neq_ops}
\end{align}
which implies that
\begin{align}
  \sum_{p\ne q} v_p v_q X_p X_q
  \EQFS - \v v\c\v v X_0 X_1
  \EQFS -\f{\v v\c\v v}{N\p{N-1}}
  \sp{\p{\col{X}}^2 - N \p{\col{X^2}}}
  \label{eq:PXXP_neq}
\end{align}
Altogether, we have that
\begin{align}
  \V_1\p{\v v}^2
  \EQFS -\f{\v v\c\v v}{N\p{N-1}} \sp{\p{\col{X}}^2 - N \p{\col{X^2}}}.
\end{align}

@@

Calculating the first-order effective Hamiltonian $H_2^{(1)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
\eqref{eq:perturbations} requires us to simplify
\begin{align}
  \P_0 \V_2 \P_0
  \EQFS \f12 \sum_{p\ne q} w_{pq} X_p Y_q
  = {N\choose 2} \bar w X_0 Y_1,
  \label{eq:PXYP_start}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, ${N\choose2}=N\p{N-1}/2$ is a binomial
coefficient, and
\begin{align}
  \bar w \equiv {N\choose 2}^{-1} \sum_{p<q} w_{pq}
\end{align}
is the mean value of the pair-wise couplings in $\V_2$.  To simplify
the operator content of \eqref{eq:PXYP_start}, we define
\begin{align}
  \col{X} \equiv \sum_p X_p,
  &&
  \col{Y} \equiv \sum_p Y_p,
  &&
  A \odot B \equiv \f12\p{AB+BA},
  &&
  \col{X\odot Y} \equiv \sum_p X_p \odot Y_p,
\end{align}
and expand
\begin{align}
  \col{X}\odot\col{Y} \EQFS \col{X\odot Y} + N\p{N-1} X_0 Y_1.
\end{align}
Substituting the product $X_0 Y_1$ back into \eqref{eq:PXYP_start}
gives us
\begin{align}
  \V_2 \EQFS \f12 \bar w \p{\col{X}\odot\col{Y} - \col{X\odot Y}}.
\end{align}

@@

Calculating the second-order effective Hamiltonian $H_2^{(2)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
\eqref{eq:perturbations} requires us to simplify
\begin{align}
  \P_0 \V_2^2 \P_0
  \EQFS \f14 \sum_{k,\ell,p,q}
  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q,
  \label{eq:PXYXYP_start}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, and the coefficients $\m w$ satisfy
\eqref{eq:cond_2} with eigenvalue $\Delta\p{\m w}\ne0$, which implies
that $\m w$ is mean-zero.  To simplify the sum in
\eqref{eq:PXYXYP_start}, we first consider the terms with
$p,q\in\set{k,\ell}$:
\begin{align}
  \sum_{\substack{k,\ell\\p,q\in\set{k,\ell}}}
  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
  \EQFS \v{\m w}\c\v{\m w} \p{X_0^2 Y_1^2 + X_0 Y_0 Y_1 X_1},
  &&
  \v{\m w} \equiv \sum_{k,\ell} w_{k\ell} \ket{k\ell}.
\end{align}
The terms of the sum in \eqref{eq:PXYXYP_start} with
$p\in\set{k,\ell}$ and $q\notin\set{k,\ell}$ or vice versa are
\begin{align}
  \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
  &\EQFS \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
  w_{k\ell} w_{k q} \p{X_0^2 Y_1 Y_2 + Y_0 X_0 X_1 Y_2}, \\
  \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
  &\EQFS \sum_{\substack{k,\ell\\p\notin\set{k,\ell}}}
  w_{k\ell} w_{pk} \p{X_0 Y_0 X_1 Y_2 + Y_0^2 X_1 X_2},
\end{align}
where
\begin{align}
  \sum_{\substack{k,\ell\\p\notin\set{k,\ell}}} w_{k\ell} w_{pk}
  = \sum_{k,\ell,p} w_{k\ell} w_{kp} - \sum_{k,\ell} w_{k\ell}^2
  = \v w\c\v w - \v{\m w}\c\v{\m w},
  &&
  \v w \equiv \sum_{k,\ell} w_{k\ell} \ket{k}.
\end{align}
Finally, the terms of the sum in \eqref{eq:PXYXYP_start} with
$p,q\notin\set{k,\ell}$ are
\begin{align}
  \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
  \EQFS \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
  w_{k\ell} w_{pq} X_0 X_1 Y_2 Y_3,
\end{align}
where, using the fact that $\m w$ is mean-zero,
\begin{align}
  \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}} w_{k\ell} w_{pq}
  = - \sum_{\substack{k,\ell\\p,q\in\set{k,\ell}}} w_{k\ell} w_{pq}
  - \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
  w_{k\ell} w_{pq}
  - \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
  w_{k\ell} w_{pq}
  = 2 \v{\m w}\c\v{\m w} - 4 \v w\c\v w.
\end{align}
Altogether, we thus far have that
\begin{multline}
  \P_0 \V_2^2 \P_0
  \EQFS -\p{\v w\c\v w - \f12 \v{\m w}\c\v{\m w}} X_0 X_1 Y_2 Y_3
  + \f14 \v{\m w}\c\v{\m w} \p{X_0^2 Y_1^2 + X_0 Y_0 Y_1 X_1} \\
  + \f14 \p{\v w\c\v w - \v{\m w}\c\v{\m w}}
  \p{X_0^2 Y_1 Y_2 + X_0 Y_0 X_1 Y_2 + Y_0 X_0 X_1 Y_2 + Y_0^2 X_1 X_2}.
\end{multline}
If $X=Y=s_\z$, where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin
operator, then
\begin{align}
  \P_0 \V_2^2 \P_0
  \EQFS - \sp{\f{\v w\c\v w-\v{\m w}\c\v{\m w}/2}{N_4}} S_\z^4
  + \sp{\f{\v w\c\v w-\v{\m w}\c\v{\m w}}{4N_2}
    + \f{\p{3N-4}\p{\v w\c\v w - \v{\m w}\c\v{\m w}/2}}{2N_4}} S_\z^2,
\end{align}
where $S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin
operator,
\begin{align}
  N_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!},
\end{align}
and we have neglected scalar terms with no physical consequence.

>>>>

We now return to the task of computing the second-order effective
Hamiltonian $H_1^{(2)}$.  Any coefficient vector $\v v$ can be
expanded into its projections $\v v_\Delta$ onto the eigenspace of
vectors satisfying \eqref{eq:cond_1} with eigenvalue $\Delta$,
i.e.~$\v v = \sum_\Delta \v v_\Delta$, which also allows us to expand
\begin{align}
  \V_1\p{\v v} = \sum_\Delta \V_1\p{\v v_\Delta},
\end{align}
where each operator $\V_1\p{\v v_\Delta}$ generates a state with
definite interaction energy $\Delta$ above that of the fully symmetric
manifold $\M_0$.  We can therefore simplify
\begin{align}
  H_1^{(2)}
  = - \P_0 \V_1\p{\v v} \E \V_1\p{\v v} \P_0
  = - \sum_{\Delta>0} \f1{\Delta}
  \P_0 \V_1\p{\v v} \P_\Delta \V_1\p{\v v} \P_0
  = - \sum_{\Delta>0} \f1{\Delta} \P_0 \V_1\p{\v v_\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \V_1\p{\v v_\Delta}^2 \P_0$ is worked out in
Appendix \ref{sec:PXXP}.  In total, we find
\begin{align}
  H_1^{(2)}
  \EQFS \f1{N\p{N-1}}
  \sum_{\Delta>0} \f{\v v_\Delta\c\v v_\Delta}{\Delta}
  \sp{\p{\col{X}}^2 - N \p{\col{X^2}}},
  \label{eq:H_1_2}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, $\v v_\Delta$ is the projection of $\v v$
onto the $\Delta$-eigenspace of the matrix $\m h-\diag\v h$, and
\begin{align}
  \col{X} \equiv \sum_p X_p,
  &&
  \col{X^2} \equiv \sum_p X_p^2.
\end{align}

@@

If we {\it construct} a perturbation $\V_2\p{\m w_\Delta}$ with
coefficients $\m w_\Delta$ that satisfy \eqref{eq:cond_2_eig} with
eigenvalue $\Delta/2$, then
\begin{align}
  H_0 \V_2\p{\m w_\Delta} \ket\psi
  = \p{E_0 + \Delta} \V_2\p{\m w_\Delta} \ket\psi,
\end{align}
which implies that $\V_2\p{\m w_\Delta}\ket\psi$ is a state of
definite energy that depends only on the coefficients $\m w_\Delta$.
As with the case of single-body perturbations, we can decompose any
matrix $\m w$ into its projections $\m w_\Delta$ onto the eigenspace
of matrices satisfying \eqref{eq:cond_2_eig} with eigenvalue
$\Delta/2$, i.e.~$\m w=\sum_\Delta\m w_\Delta$, which allows us to
expand
\begin{align}
  \V_2\p{\m w} = \sum_\Delta \V_2\p{\m w_\Delta},
\end{align}
and in turn write
\begin{align}
  H_2^{(2)} = - \P_0 \V_2\p{\m w} \E \V_2\p{\m w} \P_0
  = -\sum_{\Delta>0} \f1\Delta \P_0 \V_2\p{\m w_\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \V_2\p{\m w_\Delta}^2 \P_0$ is simplified in
Appendix \ref{sec:PXYXYP}.  If $X=Y=s_\z$, where
$s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin operator, then we can expand
\begin{align}
  H_2^{(2)} \EQFS W_4\p{\m h,\m w} S_\z^4 - W_2\p{\m h,\m w} S_\z^2,
\end{align}
where $S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin
operator, $W_k\p{\m h,\m w}$ are scalar coefficients, and we have
neglected constant terms in $H_2^{(2)}$ without physical consequence.
In order to expand the coefficients $W_k\p{\m h,\m w}$, for any matrix
$\m m$ we define the vectors
\begin{align}
  \v{\m m} \equiv \sum_{p,q} m_{pq} \ket{pq},
  &&
  \v m \equiv \sum_{p,q} m_{pq} \ket{p},
\end{align}
using which we have
\begin{align}
  W_4\p{\m h,\m w}
  \equiv \sum_\Delta \f1\Delta \f{\v w_\Delta\c\v w_\Delta
    - \v{\m w}_\Delta\c\v{\m w}_\Delta/2}{N!_4},
\end{align}
\begin{align}
  W_2\p{\m h,\m w}
  \equiv \sum_\Delta \f1{\Delta} \sp{\f{\v w_\Delta\c\v w_\Delta
      - \v{\m w}_\Delta\c\v{\m w}_\Delta}{4N!_2}
    + \f{\p{3N-4}\p{\v w_\Delta\c\v w_\Delta
        - \v{\m w}_\Delta\c\v{\m w}_\Delta/2}}{2N!_4}},
\end{align}
where
\begin{align}
  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}.
\end{align}
Note that the dependence of the coefficients of $W_k\p{\m h,\m w}$ on
the interaction matrix $\m h$ is implicit in the decomposition of the
coefficient matrix $\m w$ into components $\m w_\Delta$ as
$\m w=\sum_\Delta\m w_\Delta$, where $\m w_\Delta$ is the projection
of $\m w$ onto the $\p{\Delta/2}$-eigenspace of the operator
$\tilde{\m h}$ defined in \eqref{eq:h_super_mat}.

@@

Calculating the second-order effective Hamiltonian $H_1^{(2)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_1$ in
\eqref{eq:perturbations} requires us to simplify
\begin{align}
  \P_0 \V_1\p{\v v}^2 \P_0
  \EQFS \sum_{p,q} v_p v_q X_p X_q
  = \sum_p v_p^2 X_p^2
  + \sum_{p\ne q} v_p v_q X_p X_q,
  \label{eq:PXXP_start}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$; and the coefficients $\v v$ satisfy
\eqref{eq:cond_1} with eigenvalue $\Delta\ne0$, which implies that
$\v v$ is mean-zero.  The first sum on the right of
\eqref{eq:PXXP_start} is then
\begin{align}
  \sum_p v_p^2 X_p^2
  \EQFS \sum_p v_p^2 X_0^2
  \EQFS \f1N \v v\c\v v\, \col{X^2},
  &&
  \col{X^2} \equiv \sum_p X_p^2.
  \label{eq:PXXP_eq}
\end{align}
The second sum on the right of \eqref{eq:PXXP_start} is
\begin{align}
  \sum_{p\ne q} v_p v_q X_p X_q
  \EQFS \sum_{p\ne q} v_p v_q X_0 X_1,
  \label{eq:PXXP_neq_start}
\end{align}
where the sum over coefficients in \eqref{eq:PXXP_neq_start} can be
simplified to
\begin{align}
  \sum_{p\ne q} v_p v_q
  = \sum_{p,q} v_p v_q - \sum_{p=q} v_p v_q
  = - \v v\c\v v
\end{align}
In order to simplify the product of operators in
\eqref{eq:PXXP_neq_start}, we expand
\begin{align}
  \p{\col{X}}^2
  = \sum_p X_p^2 + \sum_{p\ne q} X_p X_q
  \EQFS \col{X^2} + N \p{N-1} X_0 X_1,
  \label{eq:PXXP_neq_ops}
\end{align}
which implies that
\begin{align}
  \sum_{p\ne q} v_p v_q X_p X_q
  \EQFS - \v v\c\v v X_0 X_1
  \EQFS -\f{\v v\c\v v}{N\p{N-1}}
  \sp{\p{\col{X}}^2 - N \p{\col{X^2}}}
  \label{eq:PXXP_neq}
\end{align}
Altogether, we have that
\begin{align}
  \P_0 \V_1\p{\v v}^2 \P_0
  \EQFS -\f{\v v\c\v v}{N\p{N-1}}
  \sp{\p{\col{X}}^2 - N \p{\col{X^2}}}.
\end{align}

@@

Calculating the first-order effective Hamiltonian $H_2^{(1)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
\eqref{eq:perturbations} requires us to simplify
\begin{align}
  \P_0 \V_2 \P_0
  \EQFS \f12 \sum_{p\ne q} w_{pq} X_p Y_q
  = {N\choose 2} \bar w X_0 Y_1,
  \label{eq:PXYP_start}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, ${N\choose2}=N\p{N-1}/2$ is a binomial
coefficient, and
\begin{align}
  \bar w \equiv {N\choose 2}^{-1} \sum_{p<q} w_{pq}
\end{align}
is the mean value of the pair-wise couplings in $\V_2$.  To simplify
the operator content of \eqref{eq:PXYP_start}, we define
\begin{align}
  \col{X} \equiv \sum_p X_p,
  &&
  \col{Y} \equiv \sum_p Y_p,
  &&
  A \odot B \equiv \f12\p{AB+BA},
  &&
  \col{X\odot Y} \equiv \sum_p X_p \odot Y_p,
\end{align}
and expand
\begin{align}
  \col{X}\odot\col{Y} \EQFS \col{X\odot Y} + N\p{N-1} X_0 Y_1.
\end{align}
Substituting the product $X_0 Y_1$ back into \eqref{eq:PXYP_start}
gives us
\begin{align}
  \P_0 \V_2 \P_0
  \EQFS \f12 \bar w \p{\col{X}\odot\col{Y} - \col{X\odot Y}}.
\end{align}

@@

Calculating the second-order effective Hamiltonian $H_2^{(2)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
\eqref{eq:perturbations} requires us to simplify
\begin{align}
  \P_0 \V_2^2 \P_0
  \EQFS \f14 \sum_{k,\ell,p,q}
  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q,
  \label{eq:PXYXYP_start}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, and the coefficients $\m w$ satisfy
\eqref{eq:cond_2} with eigenvalue $\Delta\p{\m w}\ne0$, which implies
that $\m w$ is mean-zero.  To simplify the sum in
\eqref{eq:PXYXYP_start}, we first consider the terms with
$p,q\in\set{k,\ell}$:
\begin{align}
  \sum_{\substack{k,\ell\\p,q\in\set{k,\ell}}}
  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
  \EQFS \v{\m w}\c\v{\m w} \p{X_0^2 Y_1^2 + X_0 Y_0 Y_1 X_1},
  &&
  \v{\m w} \equiv \sum_{k,\ell} w_{k\ell} \ket{k\ell}.
\end{align}
The terms of the sum in \eqref{eq:PXYXYP_start} with
$p\in\set{k,\ell}$ and $q\notin\set{k,\ell}$ or vice versa are
\begin{align}
  \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
  &\EQFS \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
  w_{k\ell} w_{k q} \p{X_0^2 Y_1 Y_2 + Y_0 X_0 X_1 Y_2}, \\
  \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
  &\EQFS \sum_{\substack{k,\ell\\p\notin\set{k,\ell}}}
  w_{k\ell} w_{pk} \p{X_0 Y_0 X_1 Y_2 + Y_0^2 X_1 X_2},
\end{align}
where
\begin{align}
  \sum_{\substack{k,\ell\\p\notin\set{k,\ell}}} w_{k\ell} w_{pk}
  = \sum_{k,\ell,p} w_{k\ell} w_{kp} - \sum_{k,\ell} w_{k\ell}^2
  = \v w\c\v w - \v{\m w}\c\v{\m w},
  &&
  \v w \equiv \sum_{k,\ell} w_{k\ell} \ket{k}.
\end{align}
Finally, the terms of the sum in \eqref{eq:PXYXYP_start} with
$p,q\notin\set{k,\ell}$ are
\begin{align}
  \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
  \EQFS \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
  w_{k\ell} w_{pq} X_0 X_1 Y_2 Y_3,
\end{align}
where, using the fact that $\m w$ is mean-zero,
\begin{align}
  \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}} w_{k\ell} w_{pq}
  = - \sum_{\substack{k,\ell\\p,q\in\set{k,\ell}}} w_{k\ell} w_{pq}
  - \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
  w_{k\ell} w_{pq}
  - \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
  w_{k\ell} w_{pq}
  = 2 \v{\m w}\c\v{\m w} - 4 \v w\c\v w.
\end{align}
Altogether, we thus have that
\begin{multline}
  \P_0 \V_2^2 \P_0
  \EQFS -\p{\v w\c\v w - \f12 \v{\m w}\c\v{\m w}} X_0 X_1 Y_2 Y_3
  + \f14 \v{\m w}\c\v{\m w} \p{X_0^2 Y_1^2 + X_0 Y_0 Y_1 X_1} \\
  + \f14 \p{\v w\c\v w - \v{\m w}\c\v{\m w}}
  \p{X_0^2 Y_1 Y_2 + X_0 Y_0 X_1 Y_2
    + Y_0 X_0 X_1 Y_2 + Y_0^2 X_1 X_2}.
  \label{eq:PXYXYP_multi_body}
\end{multline}
Denoting the collective version of any single-spin operator $Z$ by
$\col{Z}\equiv\sum_pZ_p$, defining the falling factorial
\begin{align}
  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!},
\end{align}
denoting symmetrized products by $A\odot B\equiv\p{AB+BA}/2$, and
adopting the convention that exponents are evaluated before
symmetrized products,
i.e.~$A^k\odot B\equiv\p{A^k}\odot B\ne A^{k-1}\p{A\odot B}$, the
operators in \eqref{eq:PXYXYP_multi_body} can be written in terms of
collective operators within the fully symmetric manifold through the
identities
\begin{align}
  N!_2 \p{X_0^2 Y_1^2 + X_0 Y_0 Y_1 X_1}
  \EQFS \col{X^2}\odot\col{Y^2} + \col{XY} \odot \col{YX}
  - \col{X^2\odot Y^2} - \col{\p{XY}\odot\p{YX}}
\end{align}
\begin{align}
  N!_3 &\p{X_0^2 Y_1 Y_2 + X_0 Y_0 X_1 Y_2
    + Y_0 X_0 X_1 Y_2 + Y_0^2 X_1 X_2}
  \notag \\
  &\EQFS 4\col{\p{X\odot Y}^2} + 5 \col{X^2\odot Y^2} - \f12\col{\p{XY}\odot\p{YX}}
  \notag \\
  &\quad- \f32 \col{X}\odot\col{Y^2X} - \f32 \col{XY^2}\odot\col{X}
  - \f32 \col{Y}\odot\col{X^2Y} - \f32 \col{YX^2}\odot\col{X}
  \notag \\
  &\quad- \col{X}\odot\col{YXY} - \col{Y}\odot\col{XYX}
  - 2 \col{XY}\odot\col{YX} - 2\col{X^2}\odot\col{Y^2}
  \notag \\
  &\quad+ \col{X}\,\col{Y^2}\,\col{X} + \col{Y}\,\col{X^2}\,\col{Y}
  +\col{XY}\odot\p{\col{Y}\,\col{X}} + \col{YX}\odot\p{\col{X}\,\col{Y}}.
\end{align}

[UNFINISHED]
